---
title: "Final Models"
author: "Will Brazgel"
date: "2022-11-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Libraries
```{r}
library(plyr)
library(readr)
library(dplyr)
library(stringr)
library(survival)
library(tidyverse)
```


## Baseball Reference 


### Data
```{r}
cumulative22 = read.csv("2022_cumulative.csv", encoding="UTF-8")
cumulative21 = read.csv("2021_cumulative.csv", encoding="UTF-8")
cumulative20 = read.csv("2020_cumulative.csv", encoding="UTF-8")
cumulative19 = read.csv("2019_cumulative.csv", encoding="UTF-8")
cumulative18 = read.csv("2018_cumulative.csv", encoding="UTF-8")
cumulative17 = read.csv("2017_cumulative.csv", encoding="UTF-8")
cumulative16 = read.csv("2016_cumulative.csv", encoding="UTF-8")
cumulative15 = read.csv("2015_cumulative.csv", encoding="UTF-8")
cumulative14 = read.csv("2014_cumulative.csv", encoding="UTF-8")
cumulative13 = read.csv("2013_cumulative.csv", encoding="UTF-8")
cumulative12 = read.csv("2012_cumulative.csv", encoding="UTF-8")
cumulative11 = read.csv("2011_cumulative.csv", encoding="UTF-8")
cumulative10 = read.csv("2010_cumulative.csv", encoding="UTF-8")
cumulative09 = read.csv("2009_cumulative.csv", encoding="UTF-8")
cumulative08 = read.csv("2008_cumulative.csv", encoding="UTF-8")
cumulative07 = read.csv("2007_cumulative.csv", encoding="UTF-8")
cumulative06 = read.csv("2006_cumulative.csv", encoding="UTF-8")
cumulative05 = read.csv("2005_cumulative.csv", encoding="UTF-8")
cumulative04 = read.csv("2004_cumulative.csv", encoding="UTF-8")
cumulative03 = read.csv("2003_cumulative.csv", encoding="UTF-8")
cumulative02 = read.csv("2002_cumulative.csv", encoding="UTF-8")
cumulative01 = read.csv("2001_cumulative.csv", encoding="UTF-8")
cumulative00 = read.csv("2000_cumulative.csv", encoding="UTF-8")
```

```{r}
cumulative22$Status <- 0
cumulative21$Status <- 1
cumulative20$Status <- 1
cumulative19$Status <- 1
cumulative18$Status <- 1
cumulative17$Status <- 1
cumulative16$Status <- 1
cumulative15$Status <- 1
cumulative14$Status <- 1
cumulative13$Status <- 1
cumulative12$Status <- 1
cumulative11$Status <- 1
cumulative10$Status <- 1
cumulative09$Status <- 1
cumulative08$Status <- 1
cumulative07$Status <- 1
cumulative06$Status <- 1
cumulative05$Status <- 1
cumulative04$Status <- 1
cumulative03$Status <- 1
cumulative02$Status <- 1
cumulative01$Status <- 1
cumulative00$Status <- 1
```

```{r}
joinall <- rbind(cumulative22, cumulative21, cumulative20, cumulative19, cumulative18, cumulative17, cumulative16, cumulative15, cumulative14, cumulative13, cumulative12, cumulative11, cumulative10, cumulative09, cumulative08, cumulative07, cumulative06, cumulative05, cumulative04, cumulative03, cumulative02, cumulative01, cumulative00)
```

```{r}
joinall <- joinall %>%
    distinct(Name, .keep_all = TRUE)
```


### Cleaning
```{r}
joinall <- joinall %>%
  rename(ERA_plus = `ERA.`,
         K = SO,
         K9 = SO9,
         KBB_ratio = SO.W,
         WL_ratio = W.L.) %>% 
  select(-ERA..1)
```

```{r}
joinall <- joinall %>% 
  add_row(Rk = 267, Name = "Luis Garcia HOU", Age = 25, Yrs =	3,	W = 26, L =	17, WL_ratio = .605, ERA =	3.57, G =	63, GS = 57, GF = 2, CG = 0, SHO = 0, SV =	0, IP =	325.0, H = 271, R = 136, ER =	129, HR =	43, BB = 102, IBB =	1, K = 333, 	HBP = 5, BK	= 2, WP = 18,	BF = 1325,	ERA_plus = 115,	WHIP = 1.148,	H9 = 7.5,	HR9 = 1.2, BB9 = 2.8,	K9 = 9.2,	KBB_ratio = 3.26, Name.additional = "garcilu05", Status = 0)
```

```{r}
joinall$Name = str_replace_all(joinall$Name,"\xa0"," ")

joinall$Name = str_replace(joinall$Name,"\\*","")

joinall$Name = str_replace(joinall$Name,"\\#","")

joinall$Name = str_replace(joinall$Name,"N  ez","Nunez")

joinall$Name = str_replace(joinall$Name,"F lix","Felix")

joinall$Name = str_replace(joinall$Name,"B ez","Baez")

joinall$Name = str_replace(joinall$Name,"D az","Diaz")

joinall$Name = str_replace(joinall$Name,"L pez","Lopez")

joinall$Name = str_replace(joinall$Name,"P rez","Perez")

joinall$Name = str_replace(joinall$Name,"Ru z","Ruiz")

joinall$Name = str_replace(joinall$Name,"Pe a","Pena")

joinall$Name = str_replace(joinall$Name,"Rodr guez", "Rodriguez")

joinall$Name = str_replace(joinall$Name,"Mari ez", "Marinez")

joinall$Name = str_replace(joinall$Name,"Adri n Morej n", "Adrian Morejon")

joinall$Name = str_replace_all(joinall$Name,"\\.", " ")

joinall$Name = str_remove_all(joinall$Name," ")
```

```{r}
joinall <- joinall %>% 
  filter(IP != "NA",
         IP > 10)
```


### Variable Addition
```{r}
joinall <- joinall %>% 
  mutate(SP = ifelse(GS > 5, 1,0)) %>%
  mutate(WHIP_Cat = cut( WHIP, breaks=c(0,1,1.2,1.4,1.6, 100)  , labels = c("Excellent", "Good", "Average", "Poor", "Awful"), right = FALSE))

joinall %>% 
  count(SP == 0)


```


## Baseball Savant


### Data
```{r}
savant = read.csv("savant.csv")
```


### Cleaning
```{r}
savant <- savant %>% 
  select(-X.first_name,
         -last_name,
         -X,
         -player_id)
```

```{r}
savant <- savant %>%
  rename(IP = p_formatted_ip,
         BAA = batting_avg,
         OBP = on_base_percent,
         QS = p_quality_start,
         xBAA = xba,
         xOBP = xobp,
         AVG_EV = exit_velocity_avg,
         AVG_LA = launch_angle_avg,
         Barrel = barrel_batted_rate,
         Hard_Hit = hard_hit_percent,
         FPS = f_strike_percent,
         Total_Pitches = n,
         FB_Rate = n_fastball_formatted,
         FB_MPH = fastball_avg_speed,
         FB_Spin = fastball_avg_spin,
         Breaking_Rate = n_breaking_formatted,
         Breaking_MPH = breaking_avg_speed,
         Breaking_Spin = breaking_avg_spin,
         Offspeed_Rate = n_offspeed_formatted,
         Offspeed_MPH = offspeed_avg_speed,
         Offspeed_Spin = offspeed_avg_spin,
         )
```

```{r}
savant_avg <- savant %>% 
  filter(Total_Pitches != "NA") %>%
  filter(FB_Rate != "NA") %>%
  filter(Breaking_Rate != "NA") %>%
  filter(Offspeed_Rate != "NA") %>%
  group_by(Name) %>% 
  summarise(FB_Rate = mean(FB_Rate),
            FB_MPH = mean(FB_MPH),
            FB_Spin = mean(FB_Spin),
            Breaking_Rate = mean(Breaking_Rate),
            Breaking_MPH = mean(Breaking_MPH),
            Breaking_Spin = mean(Breaking_Spin),
            Offspeed_Rate = mean(Offspeed_Rate),
            Offspeed_MPH = mean(Offspeed_MPH),
            Offspeed_Spin = mean(Offspeed_Spin),
            BAA = mean(BAA),
            OBP = mean(OBP),
            xBAA = mean(xBAA),
            xOBP = mean(xOBP),
            AVG_EV = mean(AVG_EV),
            AVG_LA = mean(AVG_LA),
            Barrel = mean(Barrel),
            Hard_Hit = mean(Hard_Hit),
            FPS = mean(FPS)
            )
```

```{r}
savant_avg$Name = str_replace(savant_avg$Name,"Luis Garcia", "Luis Garcia HOU")

savant_avg$Name = str_replace_all(savant_avg$Name,"\\.", " ")

savant_avg$Name = str_remove_all(savant_avg$Name," ")
```


## Draft 


### Data
```{r}
draft = read.csv("Draft.csv",encoding="UTF-8")
```


### Cleaning
```{r}
draft <- draft %>% 
  mutate(Name = name_first_last,
         Throws = throws,
         Overall_Pick = overall,
         Round = round_desc) %>% 
  select(Name,Overall_Pick,Throws,Round)

draft$Name = str_replace_all(draft$Name,"\\.", " ")

draft$Name = str_remove_all(draft$Name," ")
```


## Fullset
```{r}
fullset <- left_join(joinall, 
                    savant_avg,
                    by = "Name")
```


### Variable Addition
```{r}
fullset <- fullset %>%
  mutate(IP_perYear = IP/Yrs) %>%
  mutate(WHIP_Cat = cut( WHIP, breaks=c(0,1,1.2,1.4,1.6, 100)  , labels = c("Excellent", "Good", "Average", "Poor", "Awful"), right = FALSE))
```


## Fullset Draft
```{r}
fullset_draft <- inner_join(fullset, 
                    draft,
                    by = "Name")

fullset_draft <- fullset_draft %>% distinct(Name, .keep_all = TRUE)
```


### Variable Addition
```{r}
fullset_draft <- fullset_draft %>%
  mutate(Day1_Pick = ifelse(Overall_Pick < 61, 1, 0),
         Round1_Pick = ifelse(Overall_Pick < 31, 1, 0),
         First10_Pick = ifelse(Overall_Pick < 301, 1, 0))
```


## Statcast
```{r}
statcast <- inner_join(joinall,
                       savant_avg,
                       by = "Name")
```


### Variable Addition
```{r}
statcast <- statcast %>% 
  mutate(FB_Primary = ifelse(FB_Rate > 57.7, 1, 0))

statcast %>% 
  count(FB_Primary)

statcast <- statcast %>% 
  mutate(Breaking_Primary = ifelse(Breaking_Rate > 28.1, 1, 0))

statcast %>% 
  count(Breaking_Primary)

statcast <- statcast %>% 
  mutate(Drop_MPH = FB_MPH - Offspeed_MPH)

statcast <- statcast %>% 
  mutate(IP_perYear = IP/Yrs)

statcast_SP <- statcast %>% 
  filter(SP == 1)

statcast_RP <- statcast %>% 
  filter(SP == 0)
```



## Handedness
```{r}
handedness <- fullset_draft %>%
  filter(Name != 'AaronBarrett',
         Name != 'JeanMachi',
         Name != 'TimHamulack',
         Name != 'BenDiggins')
```

## Statcast TVC
```{r}
statcastTVC = read.csv("statcastTVC.csv")
```


## Final Models

### Difference Between SP and RP
```{r}
fullset %>% 
  count(SP == 1)
```

```{r}
km_position = survfit(Surv(IP,Status)~SP, conf.type = "plain", data = fullset)

plot(km_position, lty = 1:2)
```

```{r}
survdiff(Surv(IP,Status)~SP, data = fullset)
```

> Conclusion : Because our p-value is incredibly small, there is enough evidence to reject the null hypothesis and further conclude that for some time points t, there is a difference in the predicted career length of a MLB pitcher dependent on whether they are a Starting Pitcher or a Relief Pitcher.


### Statcast Variable Importance/Choosing a Parametric Distribution
```{r}
CoxSnell = function(cs,status,xlim=NULL,ylim=NULL)
{
kmcs = survfit( Surv(jitter(cs,amount=(max(cs)-min(cs))/1000),status) ~ 1 )$surv

plot( log(-log(kmcs)) ~ sort(log(cs)) ,
      xlab="log(Cox-Snell)", ylab="log(-log(S(Cox-Snell)))", xlim=xlim, ylim=ylim )

abline(0,1,col='red')
}
```

```{r}
survreg(Surv(IP,Status)~FB_Rate + FB_MPH + FB_Spin, dist = "weibull", data = statcast_SP)
```

```{r}
CS = -log( 1 - pweibull( statcast_SP$IP , shape = 1/(0.9812908), scale = exp(4.671356890 - 0.005023827*statcast_SP$FB_Rate + 0.005257949*statcast_SP$FB_MPH + 0.001011687*statcast_SP$FB_Spin) ))

CoxSnell( CS , statcast_SP$Status )
```

```{r}
survreg(Surv(IP,Status)~FB_Rate + FB_MPH + FB_Spin, dist = "lognormal", data = statcast_SP)
```

```{r}
CS = -log( 1 - plnorm( statcast_SP$IP , sdlog = 1.346779, meanlog = 0.907028291 - 0.006360853*statcast_SP$FB_Rate + 0.040217193*statcast_SP$FB_MPH + 0.001084374*statcast_SP$FB_Spin ))

CoxSnell( CS , statcast_SP$Status )
```

```{r}
AIC_w = 2*(5-(-2923.8))
AIC_w
```

```{r}
AIC_lnorm = 2*(5-(-2924.6))
AIC_lnorm
```

> For SP and FB, lognormal is preferred. These calculations(CS Resids and AIC) were done before making each parametric assumption for the following models.

```{r}
SP_FB = survreg(Surv(IP,Status)~FB_Rate + FB_MPH + FB_Spin, dist = "lognormal", data = statcast_SP)
summary(SP_FB)
```

```{r}
SP_Breaking = survreg(Surv(IP,Status)~Breaking_Rate + Breaking_MPH + Breaking_Spin, dist = "weibull", data = statcast_SP)
summary(SP_Breaking)
```

```{r}
SP_Offspeed = survreg(Surv(IP,Status)~Offspeed_Rate + Offspeed_MPH + Offspeed_Spin, dist = "weibull", data = statcast_SP)
summary(SP_Offspeed)
```

```{r}
RP_FB = survreg(Surv(IP,Status)~FB_Rate + FB_MPH + FB_Spin, dist = "weibull", data = statcast_RP)
summary(RP_FB)
```

```{r}
RP_Breaking = survreg(Surv(IP,Status)~Breaking_Rate + Breaking_MPH + Breaking_Spin, dist = "weibull", data = statcast_RP)
summary(RP_Breaking)
```

```{r}
RP_Offspeed = survreg(Surv(IP,Status)~Offspeed_Rate + Offspeed_MPH + Offspeed_Spin, dist = "weibull", data = statcast_RP)
summary(RP_Offspeed)
```


### Handedness
```{r}
hand_logn = survreg(Surv(IP,Status)~Throws , dist = "lognormal", data = handedness)
summary(hand_logn)
```

```{r}
hand_w_int = survreg(Surv(IP,Status)~Throws * SP, dist = "weibull", data = handedness)
summary(hand_w_int)
```

```{r}
coxph(Surv(IP,Status)~Throws * SP, data = handedness)
```


### Pitcher Type
```{r}
coxph(Surv(IP,Status)~ FB_Primary, data = statcast_SP)

coxph(Surv(IP,Status)~ FB_Primary, data = statcast_RP)
```

```{r}
FBprimary_SP = survreg(Surv(IP,Status)~FB_Primary, dist = "weibull", data = statcast_SP)
summary(FBprimary_SP)

FBprimary_RP = survreg(Surv(IP,Status)~FB_Primary, dist = "weibull", data = statcast_RP)
summary(FBprimary_RP)
```

```{r}
msp = survreg(Surv(IP,Status)~FB_Primary + FB_MPH, dist = "lognormal", data = statcast_SP)
summary(msp)

mrp = survreg(Surv(IP,Status)~FB_Primary + FB_MPH, dist = "lognormal", data = statcast_RP)
summary(mrp)
```

```{r}
Breaking_primary_SP = survreg(Surv(IP,Status)~Breaking_Primary, dist = "weibull", data = statcast_SP)
summary(Breaking_primary_SP)

Breaking_primary_RP = survreg(Surv(IP,Status)~Breaking_Primary, dist = "weibull", data = statcast_RP)
summary(Breaking_primary_RP)
```

```{r}
coxph(Surv(IP,Status)~K9*FB_Primary, data = statcast_SP)
```


### Draft Status
```{r}
draft_overall = survreg(Surv(IP,Status)~Overall_Pick, dist = "weibull", data = fullset_draft)
summary(draft_overall)
```

```{r}
draft_round1 = survreg(Surv(IP,Status)~Round1_Pick, dist = "lognormal", data = fullset_draft)
summary(draft_round1)

draft_day1 = survreg(Surv(IP,Status)~Day1_Pick, dist = "lognormal", data = fullset_draft)
summary(draft_day1)

draft_first10 = survreg(Surv(IP,Status)~First10_Pick, dist = "lognormal", data = fullset_draft)
summary(draft_first10)
```

```{r}
coxph(Surv(IP,Status)~Round1_Pick, data = fullset_draft)
```




```{r}
coxph(Surv(IP,Status)~Round1_Pick + strata(WHIP_Cat), data = fullset_draft)

```


```{r}
cpts  = coxph(Surv(IP, Status)~ FB_Primary + strata(WHIP_Cat), data = statcast_SP )
cpts
```

```{r}
coxph(Surv(IP, Status)~ FB_Primary + strata(WHIP_Cat), data = statcast_RP )
cptr = coxph(Surv(IP, Status)~ FB_Primary + strata(WHIP_Cat), data = statcast_RP )

```

```{r}
cox.zph(cpts)
cox.zph(cptr)
```

```{r}
plot(cox.zph(cpts), xlab = "Innings", ylab = "Starters by Pitcher Type")
plot(cox.zph(cptr), xlab = "Innings", ylab = "Relievers by Pitcher Type")
```

```{r}
kmts = survfit(Surv(IP, Status)~ FB_Primary + strata(WHIP_Cat), data = statcast_SP)
kmtr = survfit(Surv(IP, Status)~ FB_Primary + strata(WHIP_Cat), data = statcast_RP)

# plot( kmts , fun='cloglog' , mark.time=FALSE ,conf.int = FALSE,  col=1:5 )
# plot( kmtr , fun='cloglog' , mark.time=FALSE ,conf.int = FALSE, col=1:5)


```



## New Statcast with Strata


> For SP and FB, lognormal is preferred. These calculations(CS Resids and AIC) were done before making each parametric assumption for the following models.

```{r}
SP_FBs = survreg(Surv(IP,Status)~FB_Rate + FB_MPH + FB_Spin + strata(WHIP_Cat), dist = "lognormal", data = statcast_SP)
summary(SP_FBs)
```

```{r}
SP_Breakings = survreg(Surv(IP,Status)~Breaking_Rate + Breaking_MPH + Breaking_Spin + strata(WHIP_Cat), dist = "weibull", data = statcast_SP)
summary(SP_Breakings)
```

```{r}
SP_Offspeeds = survreg(Surv(IP,Status)~Offspeed_Rate + Offspeed_MPH + Offspeed_Spin + strata(WHIP_Cat), dist = "weibull", data = statcast_SP)
summary(SP_Offspeeds)
```

```{r}
RP_FBs = survreg(Surv(IP,Status)~FB_Rate + FB_MPH + FB_Spin + strata(WHIP_Cat), dist = "weibull", data = statcast_RP)
summary(RP_FBs)
```

```{r}
RP_Breakings = survreg(Surv(IP,Status)~Breaking_Rate + Breaking_MPH + Breaking_Spin + strata(WHIP_Cat), dist = "weibull", data = statcast_RP)
summary(RP_Breakings)
```

```{r}
RP_Offspeeds = survreg(Surv(IP,Status)~Offspeed_Rate + Offspeed_MPH + Offspeed_Spin + strata(WHIP_Cat), dist = "weibull", data = statcast_RP)
summary(RP_Offspeeds)
```



```{r}
draft_round1s = survreg(Surv(IP,Status)~Round1_Pick + strata(WHIP_Cat), dist = "lognormal", data = fullset_draft)
summary(draft_round1s)

draft_day1s = survreg(Surv(IP,Status)~Day1_Pick + strata(WHIP_Cat), dist = "lognormal", data = fullset_draft)
summary(draft_day1s)

draft_first10s = survreg(Surv(IP,Status)~First10_Pick + strata(WHIP_Cat), dist = "lognormal", data = fullset_draft)
summary(draft_first10s)
```

